---
title: "Project 2 - Naive Baye's"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(e1071) #Naive Baye's algorithm package
library(dplyr) #Select() and other fundamntal functions
library(gmodels) #Crostable page

Pokemon <- read.csv("./pokemon_alopez247.csv", header = TRUE) #Load data
Pokemon <- Pokemon[sample(nrow(Pokemon)),] #Randomize rows

# Select Catch_range-ranges given current max/min
min_ <- min(Pokemon$Catch_Rate)
max_ <- max(Pokemon$Catch_Rate)
diff_ <- max_-min_
Catch_range_set <- c(min_,min_+(1/6)*diff_,min_+(2/6)*diff_,min_+(3/6)*diff_,min_+(4/6)*diff_,min_+(5/6)*diff_,max_)
Pokemon$Catch_Rate_Cat <- cut(Pokemon$Catch_Rate, Catch_range_set) # Create new column

# Select Attack given current max/min
min_ <- min(Pokemon$Attack)
max_ <- max(Pokemon$Attack)
diff_ <- max_-min_
Attack_set <- c(min_,min_+(1/6)*diff_,min_+(2/6)*diff_,min_+(3/6)*diff_,min_+(4/6)*diff_,min_+(5/6)*diff_,max_)
Pokemon$Attack_Cat <- cut(Pokemon$Attack, Attack_set) # Create new column

# Select Total given current max/min
min_ <- min(Pokemon$Total)
max_ <- max(Pokemon$Total)
diff_ <- max_-min_
Total_set <- c(min_,min_+(1/6)*diff_,min_+(2/6)*diff_,min_+(3/6)*diff_,min_+(4/6)*diff_,min_+(5/6)*diff_,max_)
Pokemon$Total_Cat <- cut(Pokemon$Total, Total_set) # Create new column

Pokemon <- select(Pokemon,isLegendary,Name,Type_1,Color,hasGender,Egg_Group_1,hasMegaEvolution,Body_Style,Catch_Rate_Cat,Total_Cat)
```

--------------------------------------

\begin{center}
\textbf{Ola Tranum Arnegaard - 10/08/2019}
\end{center}

--------------------------------------

\textit{\textbf{Naive Baye's Implementation:} Please implement the Naive Baye's algorithm and submit the deliverables as specified in the project rubric document under Modules.}

# Intro
Use Naive Baye's to predict if \texttt{isLegendary} is **True** for a Pokemon given its other features.

### Preprocessing
- All Pokemon are randomized (rows)
- Only included categorical features
- Catch_Rate_Cat and Total_Cat are created from numerical to categorical variables arranged with respect to their given max and min values.
- Removed categorical features that are clearly dependent on eachother

This is what the data looks like:
```{r}
head(Pokemon,3)
```

# Implementation
### Create Model
```{r}
train_data <- Pokemon[1:675,-c(1,2)] #Train data (no names or labels)
train_labels <- Pokemon[1:675,1] #Corresponding prediction labels
Pokemon_NB_model <- naiveBayes(train_data, train_labels) #Create model
```

### Prediction

```{r}
test_data <- Pokemon[676:721,-c(1,2)] #Corresponding to the training data
test_labels <- Pokemon[676:721,1] #Actual labels
Pokemon_NB_prediction <- predict(Pokemon_NB_model, test_data) #Predicted labels

CrossTable(Pokemon_NB_prediction, test_labels,
           prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```
# Results
Because the dataset is randomized, the results are different from every compilation. However, in most examples, the predictions are close to 100% correct.
In one compilation, we predicted 5 legendary Pokemon correctly; however, we also predicted one to incorrectly be legendary (total of 6). In conclusion, we predicted 83.3% correctly for \texttt{isLegendary} **True** and correspondingly 97.6% \texttt{isLegendary} correctly to be **False**.

#### Prediction model

- Our prediction picks **True** or **False** for \texttt{isLegendary} based on which has highest \texttt{posterior probability}.
- This probability is calculated as follows $$P(\texttt{isLegendary} | feat1, feat2,...featN) = \frac{P(feat1, feat2,...featN | \texttt{isLegendary})P(\texttt{isLegendary})}{P(feat1, feat2,...featN)}$$
- Our \textit{prior and conditional probabilities} as well as our \textit{likelihood} of \texttt{isLegendary} are calculated in the model which we use in our prediction. This can be seen below:

```{r}
Pokemon_NB_model
```
### Extras
For those interested, it is also interesting to see which Pokemon that were actually and mistakenly predicted to \texttt{isLegendary} from the test data.
```{r}
Names = as.data.frame(Pokemon[676:721,2]) #Select the test data names
Names$actually = test_labels #Actual boolean isLegendary values
Names$predicted = as.data.frame(Pokemon_NB_prediction) #Predicted isLegendary
```
#### - Actually \texttt{isLegendary}:
```{r}
Names[Names$actually=='True',1] #Names that isLegendary
```
#### - Predicted \texttt{isLegendary}:
```{r}
Names[Names$predicted=='True',1] #Names predicted isLegendary
```



