---
title: "Homework 1"
output: pdf_document
---

```{r setup, include=FALSE}
library(mosaic)
library(babynames)
knitr::opts_chunk$set(echo = TRUE)
```

--------------------------------------
\begin{center}
\textbf{Ola Tranum Arnegaard - 10/08/2019}
\end{center}

# Exercise 3.1
### Intro
Using the famous Galton data set from the mosaicData package:

```{r}
head(Galton)
```
\newpage
#### 1. Create a scatterplot of each person's height against their father's height

```{r}
plot1 <- ggplot(Galton, aes(x = Galton$father, y = Galton$height)) + geom_point()
plot1 <- plot1 + xlab("Father height") + ylab("Child")
plot1
```
\newpage

#### 2. Separate your plot into facets by sex

```{r}
plot1 <- plot1 + facet_wrap(Galton$sex,nrow=1,ncol=2)
plot1
```
\newpage

#### 3. Add regression lines to all of your facets

```{r}
plot1 <- plot1 + geom_smooth(method = 'auto')
plot1
```

\newpage
# Exercise 3.2
### Intro
Using the RailTrail data set from the mosaicData package:

```{r}
head(RailTrail)
```
\newpage

#### 1. Create a scatterplot of the number of crossings per day volume against the high temperature that day

```{r}
plot2 <- ggplot(RailTrail, aes(x = RailTrail$hightemp, y = RailTrail$volume)) + geom_point()
plot2 <- plot2 + xlab("Highest temperature") + ylab("Number of crossings")
plot2
```
\newpage

#### 2. Separate your plot into facets by \texttt{weekday}

```{r}
plot2 <- plot2 + facet_wrap(RailTrail$weekday,nrow=1,ncol=2)
plot2
```
\newpage

#### 3. Add regression lines to all of your facets

```{r}
plot2 <- plot2 + geom_smooth(method = 'auto')
plot2
```
\newpage

# Exercise 3.3

Angelica Schuyler Church (1756{1814) was the daughter of New York Governer Philip Schuyler and sister of Elizabeth Schuyler Hamilton. Angelica, New York was named after her. Generate a plot of the reported **proportion of babies born with the name Angelica over time and interpret the figure.**

```{r}
head(babynames)
```

```{r}
df <- as.data.frame(c(babynames[babynames$name=='Angelica','year'],babynames[babynames$name=='Angelica','prop']))

plot3 <- ggplot(df, aes(x = df$year, y = df$prop)) + geom_point()
plot3 <- plot3 + xlab("Year") + ylab("Proportion of babies born as Angelica")
plot3
```
#### Result
You see that it is a clear rise in the proportion of babies named Angelica throughout the second half of the ninenteenth century. However, throughout the 2000's the named have experienced a downfall.
\newpage

# Exercise 3.4

### Intro
The following questions use the Marriage data set from the mosaicData package.

```{r}
head(Marriage, 4)
```
\newpage

#### 1. Create an informative and meaningful data graphic.

```{r}
plot4 <- ggplot(Marriage, aes(x = Marriage$age, y = Marriage$prevcount)) + geom_point(aes(color = prevconc)) + labs(color = "Reason for \nending marriage") + facet_wrap(Marriage$person,nrow=1,ncol=2)
plot4 <- plot4 + xlab("Age") + ylab("Number of previous marriages")
plot4
```
\newpage

#### 2. Identify each of the visual cues that you are using, and describe how they are related to each variable.
- Colors are used as a visual for the feature \texttt{reason for ending marriage}
- Seperation of plot is used to seperate \texttt{brides} and \texttt{grooms}
- Scatter plot is used to compare \texttt{previous marriages} and \texttt{age}

#### 3. Create a data graphic with at least five variables (either quantitative or categorical). For the purposes of this exercise, do not worry about making your visualization meaningful|just try to encode five variables into one plot.

```{r}
plot4 + geom_point(size = 1.5, aes(color = prevconc,shape = race))
```


